Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2021-08-17 article/draft.tex
rewrite a bit lack of data part

In this article we have assessed three broad categories of interventions to tackle misinformation by Facebook, Twitter and YouTube: suspension of accounts, introduction of notices and flags and reducing the visibility of content. We have provided a methodology that can be useful for third-party monitoring. The code needed to collect the data and create the figures is available on the following public GitHub repository: \href{https://github.com/medialab/truth-and-trust-online-2021/}{https://github.com/medialab/truth-and-trust-online-2021/}, and we hope it can be a useful ressource for researchers, NGOs and journalists addressing online misinformation.

% For a detailed list of interventions of multiple platforms, we invite the reader to navigate through the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab} and to read the related article.

@@

In the light of the present research, we discuss in this section data and policy related recommandations. Mainstream platforms such as Facebook, Twitter and Youtube are increasingly working towards higher data transparency (see links that redirect to the respective transparency centers in Appendix \ref{links}). Yet the present research sheds light on the lack and the splintering of data, which impede the academic community from studying successfully online misinformation and assessing the impact of platforms' interventions regarding this phenomenon. Hence, few comments are in order regarding the lack of data, the lack of transparency vis-\`{a}-vis misinformation related interventions, the splintering of data and the lack of interventions designed specifically for misinformation.

@@

{\bf Lack of data.} First, when Facebook, Twitter or YouTube intervene by suspending an account, a page or a channel, the data disappears. It can no longer be accessed via scraping or via the official APIs (Twitter, YouTube, CrowdTangle) for the purposes of scientific research. This hampers the investigation of how platforms moderate content. To further illustrate this point, the original list of YouTube videos summarized in table \ref{tab2} with failed fact-checks, had over $200$ YouTube videos available in March $2021$. However, by June $2021$, thirty videos had disappeared both from YouTube and the YouTube API for policy violation. Similarly, in a related research project (see Thero and Vincent (2020)~\cite{therovincent}) we searched on CrowdTangle for $94$ public accounts sharing specific content associated with misinformation in November $2020$. We tried to collect the posts from those $94$ pages in January $2021$, to only discover that $11$ pages had disappeared from the CrowdTangle API.

@@

Aside from the lack of data due to suspensions, some essential metrics for the study of misinformation are not accessible via main APIs. In particular, the $reach$ of posts on Facebook, which is the number of people who actually see a given post, cannot be recovered from the CrowdTangle API. Hence, researchers can only build proxies for this metric based on available data (see section \ref{reduce_fb} for an example). Furthermore, to the best of our knowledge, fields related to flags, notices and information panels are not available on the official APIs; with the exception of the {\it withheld}  and {\it possibly sensitive content} interstitials in the Twitter API v2. For the purposes of the present research we thus had to scrape this information from Facebook, Twitter and YouTube. Again, this disrupts the monitoring of platforms' interventions concerning misinformation and the assessment of their impact. We refer the reader to the insightful commentary by Pasquetto, Swire-Thompson, Amazeen et al. (2020)~\cite{pasquetto} and the references therein, for a discussion on partial or biased social media data.

%{\color{red} Misinformation review REF + others to show that we are not the only ones to say that}.
\smallskip

{\bf Lack of transparency.} Second, mainstream platforms generally lack transparency regarding their misinformation related interventions.  The broad lines of policies regarding misinformation are communicated via blog posts and are not typically part of the set of rules or community guidelines of mainstream platforms (see Appendix \ref{links}). Yet, when navigating through the categories of policy areas, to the best of our knowledge, misinformation is absent from the data available on the transparency centers of Facebook, Twitter and YouTube. To illustrate, the $2021$ $Q1$ {\it Community Standards Enforcement Report} of Facebook, provides the volume of actioned content only regarding the following policy areas: (1) Adult Nudity \& Sexual Activity, (2) Bullying \& Harassement, (3) Child Nudity \& Sexual Exploitation, (4) Dangerous Orgs: Organized Hate, (5) Dangerous Orgs: Terrorism, (6) Fake Accounts, (7) Hate Speech, (8) Regulated Good: Drugs, (9) Regulated Good: Firearms, Spam, (10) Suicide and Self-injury, (11) Violent \& Graphic Content. More precisely, platforms' official communication about misinformation interventions is scarce and the academic community, NGOs and data journalists, usually discover interventions related to misinformation via monitoring social media accounts related to domain names with several failed fact-checks or via articles in News outlets. For example, while conducting the present research, it was brought to our attention that during our study, the Twitter account related to the domain name $globalresearch.ca$ was suspended. But, there was no official communication or blog post by Twitter Safety team regarding this intervention (see section \ref{globalresearch}).

@@

{\bf Scattered data.} Third, to study online misinformation, researchers, NGOs and data journalists turn to multiple APIs besides the official ones managed by the mainstream platforms themselves (e.g. BuzzSumo). They do so, because multiple data sources can be complementary and no specific API has complete data. {\color{pink} To be drafted by HT:  vs. scraping, cite the article ~\cite{api} about APIs, article du NYTimes sur CT, APIs qui ouvrent et qui ferment, limites de l'API de Twitter entre la V1 et la V2.}

%There are other ways to collect data from platforms, and besides Buzzsumo, other API are also aggregating data from multiple social platforms. For example Newsguard, blablabla... In this tweet, we can see the interface of XXX being used in this study from a data journalist: \href{https://twitter.com/Shayan86/status/1394742784683298818/photo/2}{ex}

{\bf Lack of misinformation specific interventions.} It is likely that most interventions related to misinformation correspond to interventions that were designed for other policy areas or aim violations that fall into other categories cited in the platforms' guidelines. To illustrate, YouTube explicitly states\footnote{See {\it What policies exist to fight misinformation on Youtube?} \href{https://www.youtube.com/intl/en\_us/howyoutubeworks/our-commitments/fighting-misinformation/\#policies}{youtube.com/intl/en\_us/howyoutubeworks/our-commitments/fighting-misinformation/\#policies}.} that: {\it Several policies in our Community Guidelines are directly applicable to misinformation}; such as deceptive practices, impersonation and hate speech policy. Similarly, Facebook took down four pages related to InfoWars and cited as a reason the repeated violations of Community Standards and not false news.\footnote{An article published \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{on Facebook Newsroom (2018)} states that : ``All four Pages have been unpublished for repeated violations of Community Standards and accumulating too many strikes. While much of the discussion around Infowars has been related to false news, which is a serious issue that we are working to address by demoting links marked wrong by fact checkers and suggesting additional content, none of the violations that spurred today’s removals were related to this''.}

>>>>

In this article we have assessed three broad categories of interventions to tackle misinformation by Facebook, Twitter and YouTube: suspension of accounts, introduction of notices and flags and reducing the visibility of content.
\footnote{For a detailed list of interventions of multiple platforms, we invite the reader to navigate through the following \href{https://airtable.com/shrO0ooI9WSEfIUhb/tblAWQwFOiihKdQjm/viwZLAOzLK1NQ0c2n?blocks=hide}{airtable} compiled by Slatz and Leibowicz (2021)~\cite{niemanlab}.}
Furthermore, we have provided a methodology that can be useful for third-party monitoring. 
The code needed to collect the data and create the figures is available on the following public GitHub repository: \href{https://github.com/medialab/truth-and-trust-online-2021/}{https://github.com/medialab/truth-and-trust-online-2021/}, and we hope it can be a useful ressource for researchers, NGOs and journalists addressing online misinformation.

@@

Mainstream platforms such as Facebook, Twitter and Youtube are increasingly working towards higher data transparency (see links that redirect to the respective transparency centers in Appendix \ref{links}). 
Yet the present research sheds light on the restrictions to access data and its limits, which impede the academic community from studying successfully online misinformation and assessing the impact of platforms' interventions regarding this phenomenon. 
Hence, few comments are in order regarding the limitations to access data, the lack of transparency vis-\`{a}-vis misinformation related interventions and the lack of interventions designed specifically for misinformation.

@@

{\bf A need for more complete data.} When Facebook, Twitter or YouTube intervene by suspending an account, the data can no longer be accessed via scraping or via the official APIs (Twitter, YouTube, CrowdTangle) for the purposes of scientific research. 
This hampers the investigation of how platforms moderate content. 
To further illustrate this point, the original list of YouTube videos with failed fact-checks (summarized in Table \ref{tab2}), had over $200$ YouTube videos available in March $2021$. 
However, by June $2021$, thirty videos had disappeared from the YouTube API for policy violation. 
Similarly, in a related research project (see Thero and Vincent (2020)~\cite{therovincent}) we searched on CrowdTangle for $94$ public accounts sharing specific content associated with misinformation in November $2020$. 
We tried to collect the posts from those $94$ pages in January $2021$, to only discover that $11$ pages had disappeared from the CrowdTangle API.
It is of course not surprising that plaforms are regulating their content, sometimes by removing it, but we would argue that it could simplify the researchers' work if the deleted data could still be accessible in the APIs (Pasquetto, Swire-Thompson, Amazeen et al. (2020)~\cite{pasquetto}).

@@

Aside from the disappearance of data due to content moderation, some essential metrics for the study of misinformation are not accessible via main APIs. 
In particular, the $reach$ of posts on Facebook, which is the number of people who actually see a given post, cannot be recovered from the CrowdTangle API (REF). 
Hence, researchers can only build proxies for this metric based on available data (see section \ref{reduce_fb} for an example). 
Furthermore, to the best of our knowledge, fields related to flags, notices and information panels are not available on the official APIs, with the exception of the {\it withheld}  and {\it possibly sensitive content} interstitials in the Twitter API v2. 
For the purposes of the present research we thus had to scrape this information from Facebook, Twitter and YouTube. 
Again, this disrupts the monitoring of the platforms' interventions and the assessment of their impact (Pasquetto, Swire-Thompson, Amazeen et al. (2020)~\cite{pasquetto}).

{\color{brown} A last concern concerns the increasing restrictions to access data via APIs.

1/ Restrictions to the Open Graph API with the Cambridge Analytica scandal.

2/ Now CT exists but doubt were recently raised that this API will still be accessible for researchers and journalists.

3/ Still access to the "official" APIs for Twitter and YouTube, although the access to Twitter API is very hard to justify if we say we are researchers.

4/ Increasing volume restrictions on the Twitter API between V1 and V2.

It should be noted that there are other ways to collect data than directly from the platforms' APIs. Indeed, other APIs are also aggregating data from multiple social platforms as there is a need for social marketing. CrowdTangle was actually first an independent social listening tool, before being bought and included within the Facebook organization. For example Newsguard, Buzzsumo, Linkfluence... Problem: often it is not clear how the data are collected, and at which frequency (more reliable to go to the "source", i.e. the official API).}

{\bf A need for more transparency.} Second, mainstream platforms generally lack transparency regarding their misinformation related interventions.  The broad lines of policies regarding misinformation are communicated via blog posts and are not typically part of the set of rules or community guidelines of mainstream platforms (see Appendix \ref{links}). Yet, when navigating through the categories of policy areas, to the best of our knowledge, misinformation is absent from the data available on the transparency centers of Facebook, Twitter and YouTube. To illustrate, the $2021$ $Q1$ {\it Community Standards Enforcement Report} of Facebook, provides the volume of actioned content only regarding the following policy areas: (1) Adult Nudity \& Sexual Activity, (2) Bullying \& Harassement, (3) Child Nudity \& Sexual Exploitation, (4) Dangerous Orgs: Organized Hate, (5) Dangerous Orgs: Terrorism, (6) Fake Accounts, (7) Hate Speech, (8) Regulated Good: Drugs, (9) Regulated Good: Firearms, Spam, (10) Suicide and Self-injury, (11) Violent \& Graphic Content. More precisely, platforms' official communication about misinformation interventions is scarce and the academic community, NGOs and data journalists, usually discover interventions related to misinformation via monitoring social media accounts related to domain names with several failed fact-checks or via articles in News outlets. For example, while conducting the present research, it was brought to our attention that during our study, the Twitter account related to the domain name $globalresearch.ca$ was suspended. But, there was no official communication or blog post by Twitter Safety team regarding this intervention (see section \ref{globalresearch}).

@@

{\bf A need for a global and coherent misinformation policy.} It is likely that most interventions related to misinformation correspond to interventions that were designed for other policy areas or aim violations that fall into other categories cited in the platforms' guidelines. To illustrate, YouTube explicitly states\footnote{See {\it What policies exist to fight misinformation on Youtube?} \href{https://www.youtube.com/intl/en\_us/howyoutubeworks/our-commitments/fighting-misinformation/\#policies}{youtube.com/intl/en\_us/howyoutubeworks/our-commitments/fighting-misinformation/\#policies}.} that: {\it Several policies in our Community Guidelines are directly applicable to misinformation}; such as deceptive practices, impersonation and hate speech policy. Similarly, Facebook took down four pages related to InfoWars and cited as a reason the repeated violations of Community Standards and not false news.\footnote{An article published \href{https://about.fb.com/news/2018/08/enforcing-our-community-standards/}{on Facebook Newsroom (2018)} states that : ``All four Pages have been unpublished for repeated violations of Community Standards and accumulating too many strikes. While much of the discussion around Infowars has been related to false news, which is a serious issue that we are working to address by demoting links marked wrong by fact checkers and suggesting additional content, none of the violations that spurred today’s removals were related to this''.}
